---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# exploreSE

<!-- badges: start -->
<!-- badges: end -->

The goal of exploreSE is to create and maintian a shiny app for transcriptomic results presentation. It heavily relies on the [SummarizedExperiment](https://www.bioconductor.org/packages/release/bioc/html/SummarizedExperiment.html) and [DeeDeeExperiment](https://bioconductor.org/packages//release/bioc/html/DeeDeeExperiment.html) packages on bioconductor. 

## Installation

You can install the development version of exploreSE from [GitHub](https://github.com/) with:

``` r
# install.packages("pak")
pak::pak("jaspitzer/exploreSE")
```

## Example

Let's say you've loaded in your data and created two separate models:

```{r example}
library(exploreSE)
library(airway)
library(DeeDeeExperiment)
library(DESeq2)
data(airway)
```

```{r}
airway <- DESeqDataSet(airway, design = ~dex)
airway <- DESeq(airway)
baseline <- results(airway)

airway <- DESeqDataSet(airway, design = ~dex + cell)
airway <- DESeq(airway)
cell_controlled <- results(airway)
```

These results can be added to the DeeDeeExperiment
```{r}
airway <- addDEA(DeeDeeExperiment(airway), baseline)
airway <- addDEA(airway, cell_controlled)
```

```{r}
airway
```

```{r}
airway <- get.gos(obj = airway, NAME = "baseline", gene_type = "ENSEMBL")
airway <- get.gos(obj = airway, NAME = "cell_controlled", gene_type = "ENSEMBL")
```

```{r}
airway
```

With the enrichments added, we can now call our shiny app for exploration! (Note: this is a work in progress, I still need to figure out how to properly show this in the vignette)
```{r}
app <- exploreSE()
if(interactive()){
  shiny::runApp(app, port = 1234)
}
```

